---
title: "Congo Maternal Stress Infant Gut Microbiome QC"
author: "Christopher Dutton"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    code_folding: hide
  
---

This file starts with the fastq files from the PACBIO run and runs through the dada2 workflow to get a phyloseq object at the end. Two cells were used with all samples ran once on each cell. Fastq files from the different cells are them combined after error learning for the individual runs. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## 100gb ram and 12 cores on hipergator
```

```{r warning=FALSE, message=FALSE}
## https://benjjneb.github.io/LRASManuscript/LRASms_fecal.html

library(dada2)
library(Biostrings)
library(ShortRead)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(phyloseq)
library(gtools)

```

# Cell 1

```{r warning=FALSE, message=FALSE}
## path1 <- "/blue/mulligan/duttonc/Congo/Cell1" 
path1 <- "C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell1" 

fns1 <- list.files(path1, pattern="fastq", full.names=TRUE)
F27 <- "AGRGTTYGATYMTGGCTCAG"
R1492 <- "RGYTACCTTGTTACGACTT"
rc <- dada2:::rc
theme_set(theme_bw())

```

```{r warning=FALSE, message=FALSE}

if (file.exists("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell1/cell1_output.Rdata")){
  load("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell1/cell1_output.Rdata")
} else{}

## remove primers
nops1 <- file.path(path1, "noprimers", basename(fns1))
if (exists("prim1")){
  print("cached file loaded: prim1")
} else{
  prim1 <- removePrimers(fns1, nops1, primer.fwd=F27, primer.rev=dada2:::rc(R1492), orient=TRUE)
}

## filter
if (exists("lens.fn1")){
    print("cached file loaded: lens.fn1")
  } else{
  lens.fn1 <- lapply(nops1, function(fn) nchar(getSequences(fn)))
}

lens <- do.call(c, lens.fn1)
hist(lens, 1000)

filts1 <- file.path(path1, "noprimers", "filtered", basename(fns1))

if (exists("track1")){
  print("cached file loaded: track1")
} else{
  track1 <- filterAndTrim(nops1, filts1, minQ=3, minLen=1000, maxLen=1600, maxN=0, rm.phix=FALSE, maxEE=2)
}

track1

## dada2
if (exists("drp1")){
  print("cached file loaded: drp1")
} else{
  drp1 <- derepFastq(filts1, verbose=TRUE)
}


## Learn errors
if (exists("err1")){
  print("cached file loaded: err1")
} else{
  err1 <- learnErrors(drp1, errorEstimationFunction=PacBioErrfun, BAND_SIZE=32, multithread=TRUE)
}

## Plot errors
plotErrors(err1)

## Denoise

if (exists("dd1")){
  print("cached file loaded: dd1")
} else{
  dd1 <- dada(drp1, err=err1, BAND_SIZE=32, multithread=TRUE)
}

cbind(ccs=prim1[,1], primers=prim1[,2], filtered=track1[,2], denoised=sapply(dd1, function(x) sum(x$denoised)))

## Sequence table

if (exists("st1")){
  print("cached file loaded: st1")
} else{
  st1 <- makeSequenceTable(dd1); dim(st1)
  
  save(lens.fn1, dd1, drp1, err1, prim1, track1, st1, file="C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell1/cell1_output.Rdata")
}

```

# Cell 2

```{r warning=FALSE, message=FALSE}

## CELL 2
if (file.exists("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell2/cell2_output.Rdata")){
  load("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell2/cell2_output.Rdata")
} else{}

##path2 <- "/blue/mulligan/duttonc/Congo/Cell2"
path2 <- "C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell2" 

fns2 <- list.files(path2, pattern="fastq", full.names=TRUE)

## remove primers
nops2 <- file.path(path2, "noprimers", basename(fns2))
if (exists("prim2")){
  print("cached file loaded: prim2")
} else{
  prim2 <- removePrimers(fns2, nops2, primer.fwd=F27, primer.rev=dada2:::rc(R1492), orient=TRUE)
}

## filter
if (exists("lens.fn2")){
  print("cached file loaded: lens.fn2")
} else{
  lens.fn2 <- lapply(nops2, function(fn) nchar(getSequences(fn)))
}

lens <- do.call(c, lens.fn2)
hist(lens, 1000)

filts2 <- file.path(path2, "noprimers", "filtered", basename(fns2))

if (exists("track2")){
  print("cached file loaded: track2")
} else{
  track2 <- filterAndTrim(nops2, filts2, minQ=3, minLen=1000, maxLen=1600, maxN=0, rm.phix=FALSE, maxEE=2)
}

track2

## dada2
if (exists("drp2")){
  print("cached file loaded: drp2")
} else{
  drp2 <- derepFastq(filts2, verbose=TRUE)
}

## Learn errors
if (exists("err2")){
  print("cached file loaded: err2")
} else{
  err2 <- learnErrors(drp2, errorEstimationFunction=PacBioErrfun, BAND_SIZE=32, multithread=TRUE)
}

## Plot errors
plotErrors(err2)

## Denoise
if (exists("dd2")){
  print("cached file loaded: dd2")
} else{
  dd2 <- dada(drp2, err=err2, BAND_SIZE=32, multithread=TRUE)
}

cbind(ccs=prim2[,1], primers=prim2[,2], filtered=track2[,2], denoised=sapply(dd2, function(x) sum(x$denoised)))

## Sequence table
if (exists("st2")){
  print("cached file loaded: st2")
} else{
  st2 <- makeSequenceTable(dd2); dim(st2)
  
    save(lens.fn2, dd2, drp2, err2, prim2, track2, st2, file="C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/Cell2/cell2_output.Rdata")
}
```

# Combine sequence tables

```{r warning=FALSE, message=FALSE}

## due to error made at sequencing facility, need to change the name of one of the samples in each sequence table
rownames(st1) <- gsub("G13_3M_2", "G37_3M", rownames(st1))
rownames(st2) <- gsub("G13_3M_2", "G37_3M", rownames(st2))

## mislabeled sample from during collection
rownames(st1) <- gsub("G7_MV2", "G7_MV1", rownames(st1))
rownames(st2) <- gsub("G7_MV2", "G7_MV1", rownames(st2))

## combine the two sequence tables together by summing them. 
stcomb <- mergeSequenceTables(st1, st2, repeats="sum")

## check rownames, should still be 120 samples
rownames(stcomb)

```

# Assign Taxonomy

```{r warning=FALSE, message=FALSE}
## Assign taxonomy, download appropriate reference database from here - https://benjjneb.github.io/dada2/training.html

## Please note, assignTaxonomy with the species training set is recommended over the other options.
## https://github.com/benjjneb/dada2/issues/1319#issuecomment-820659005

## Combine
if (file.exists("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/cellcombined_output.Rdata")){
  load("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/cellcombined_output.Rdata")
} else{}

if (exists("tax")){
  print("cached file loaded: tax")
} else{
  tax <- assignTaxonomy(stcomb, "C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/silva_nr99_v138.1_wSpecies_train_set.fa.gz", multithread=TRUE) # Slowest part
  tax[,"Genus"] <- gsub("Escherichia-Shigella", "Escherichia", tax[,"Genus"]) # Reformat to be compatible with other data sources
}

head(unname(tax))

if (exists("bim")){
  print("cached file loaded: bim")
} else{
  bim <- isBimeraDenovo(stcomb, minFoldParentOverAbundance=3.5, multithread=FALSE)

}
```

What proportion are chimeras?

```{r warning=FALSE, message=FALSE}
## Check Chimeras
if (file.exists("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/cellcombined_output.Rdata")){
  load("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/cellcombined_output.Rdata")
} else{

  bim <- isBimeraDenovo(stcomb, minFoldParentOverAbundance=3.5, multithread=FALSE)

}

table(bim)
sum(stcomb[,bim])/sum(stcomb)

## Extract Sample Names
sample.names <- sapply(strsplit(fns2, "Cell2/"), function(x) paste(x[2]))
sample.names <- sapply(strsplit(sample.names, "-"), function(x) paste(x[1]))
rownames(stcomb) <- sample.names
sample.names

## remove chimeras
if (exists("seqtab")){
  print("cached file loaded: seqtab")
} else{
  seqtab <- removeBimeraDenovo(stcomb, method="consensus", multithread=FALSE)
  
  save(bim, seqtab, tax, file=("C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/cellcombined_output.Rdata"))
}

```

```{r warning=FALSE, message=FALSE}

## construct the phyloseq object 

## import sample data
sample_data <- read.csv(file="C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/CombinedMetadata_v8.csv", header=TRUE, sep=",")
rownames(sample_data) <- sample_data$sampleId

otuforsamples <- otu_table(seqtab, taxa_are_rows = FALSE)

sample_names(otuforsamples) <- paste(sample_names(otuforsamples), "_2", sep="")

dataforsamples<-sample_data(sample_data)

psCongo2 <- phyloseq(otuforsamples, dataforsamples, tax_table(tax))

saveRDS(psCongo2, "C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/psCongo2_V3.rds") 

psCongo2

```

Import the first round Phyloseq Object and combine it with the new object to conduct analyses over the full year time period.

```{r}

metadata <- read.csv(file="CombinedMetadata_V8.csv", header = TRUE, sep=",")

# ## fix names, append with _2
# testdf <- sample_data(psCongo2)
# testdf$newname <- paste(testdf$sampleId, "_2", sep="")
# sample_names(psCongo2) <- testdf$newname
# psCongo2

psCongo_V4 <- readRDS("psCongo_V4.rds") ## this is the first round phyloseq object

## need to correct a naming error from this original object
sample_names(psCongo_V4)[sample_names(psCongo_V4) == "G7_MV2"] <- "G7_MV1"

psCongo_V4

sampletable <- sample_data(metadata)

rownames(sampletable) <- sampletable$sampleId

pscongotax <- tax_table(psCongo_V4)
pscongootu <- otu_table(psCongo_V4)

psCongo1 <- phyloseq(pscongootu, sampletable, pscongotax)

psCongo1
psCongo2

combinedcongo <- merge_phyloseq(psCongo2, psCongo1)
combinedcongo

saveRDS(combinedcongo, "C:/Users/cldut/OneDrive - University of Florida/Dutton_Lab/Congo/SecondRound/psCombinedCongo_V5.rds") 

```


```{r}
sessionInfo()

```

